---
output: pdf_document
---
```{r include=FALSE}
library(dplyr)

downloadPriceDF <- function(stock, start = 2010, nyears = 1) {
    require(BatchGetSymbols)
    startdate <- paste(start, '01', '01', sep = '-')
    enddate <- paste(start+nyears, '01', '01', sep = '-')
    out <- BatchGetSymbols(tickers = stock, first.date = startdate, last.date = enddate, do.cache = FALSE)
    cat('\n')
    if(out$df.control$download.status != 'OK') stop(paste0('something went wrong downloading ', stock, 'prices'))
    stockDF <- data.frame(date = as.character(out$df.tickers$ref.date),
                          price = out$df.tickers$price.adjusted)
    return(stockDF)
} #uses Yahoo finance
downloadStockPairDF <- function(stock1, stock2, start = 2010, nyears = 1) {
  stock1df <- downloadPriceDF(stock1, start = start, nyears = nyears)
  stock2df <- downloadPriceDF(stock2, start = start, nyears = nyears)
  if (!identical(stock1df$date, stock2df$date)) {
    stop('Dates are wack')
    #rid <- which(stock1df$date != stock2df$date)
    #stock1df <- stock1df[-rid, ]
    #stock2df <- stock2df[-rid, ]
  }
  df <- data.frame(stock1 = stock1df$price,
                   stock2 = stock2df$price,
                   ratio = stock1df$price/stock2df$price)
  return(df)
}

plotStocks <- function(stockPairDF) {
  max <- max(max(stockPairDF$stock1), max(stockPairDF$stock2))
  plot(x = 1:length(stockPairDF$stock1), y = stockPairDF$stock1, type = 'l', col = 'red',
       ylim = c(0, max), xlab = '', ylab = 'Price')
  lines(x = 1:length(stockPairDF$stock1), y = stockPairDF$stock2, col = 'blue')
  legend('topleft', legend = c('stock1', 'stock2'), col = c('red', 'blue'), lty = 1, cex = .8)
}
plotRatio <- function(stockDF, k=1) {
  m <- mean(stockDF$ratio); s <- sd(stockDF$ratio)
  plot(x=1:length(stockDF$ratio), y=stockDF$ratio,
       type='l',
       xlab='', ylab=paste('Ratio (k=',k,')', sep = ''))
  abline(h=c(m-k*s,m,m+k*s), lty=c(2,1,2))
}
findPositions <- function(ratio, m, s, k = 1) {
  len <- length(ratio)
  toReturn <- list(); lind <- 1
  finger <- 1
  openquestionmark <- FALSE
  
  while (finger < len) {

    if (!openquestionmark) { #means we have to find the first place to open a position
      highones <- which(ratio[finger:len]>(m+s*k))
      lowones <- which(ratio[finger:len]<(m-s*k))
      if (length(highones)>0 | length(lowones)>0) {
       start <- min(highones[1], lowones[1], na.rm = TRUE) + finger - 1
       highorlow <- ifelse(identical(start, highones[1]+finger-1),
                           1, -1)
       finger <- start
       openquestionmark <- TRUE
      }
      else finger <- len
    }
    else { #we need to find where to close the position
      close <- ifelse(highorlow == 1,
                      which(ratio[finger:len]<m)[1],
                      which(ratio[finger:len]>m)[1])
      if (is.na(close)) { #means there is no good close; must close on last day
        close <- len
      }
      else close <- close+finger-1
      
      finger <- close
      toReturn[[lind]] <- c(start, close, highorlow)
      lind <- lind + 1
      openquestionmark <- FALSE
    }
       
  }#end of while loop
  return(toReturn)
}
addPositions <- function(ratio, positions) {
  for (p in positions) {
    points(x=p[1],y=ratio[p[1]],col='green',pch=19)
    points(x=p[2],y=ratio[p[2]],col='red',pch=19)
  }
}
positionProfit <- function(stocksDF, positions, net = TRUE) {
  if (length(positions) == 0) return(0)
  cashcash <- numeric(length(positions))
  i <- 1
  for (p in positions) {
    shares1 <- 1/stocksDF$stock1[p[1]]
    shares2 <- 1/stocksDF$stock2[p[1]]
    profit1 <- p[3]*-1 * shares1 * stocksDF$stock1[p[2]]
    profit2 <- p[3] * shares2 * stocksDF$stock2[p[2]]
    fees <- 0.003 * (1+1+abs(profit1)+abs(profit2))
    cashcash[i] <- profit1+profit2-fees
    i <- i+1
  }
  if (net) return(sum(cashcash))
  return(cashcash)
}
findOptimalK <- function(stocksDF, plot = FALSE) {
  ratio <- stocksDF$ratio; m <- mean(stocksDF$ratio); s <- sd(stocksDF$ratio)
  kmax <- max(abs(ratio - m))/s
  kvalues <- seq(0, kmax, length = 100)
  kprof <- sapply(kvalues, function(x) positionProfit(stocksDF = stocksDF,
                                                      findPositions(ratio, m, s, k=x)))
  ind <- which(kprof == max(kprof))[1]
  bestk <- kvalues[ind]
  
  if (plot) {
    plot(x=kvalues, y=kprof, type='p',
         xlab='k value', ylab='Profit')
    points(x=bestk, y=kprof[ind], pch=19, col='red')
  }
  
  return(bestk)
}
evaluatePairsTrading <- function(stocksDF, trainingFrac = 0.5, plot = FALSE) {
  cutoff <- ceiling(nrow(stocksDF)*trainingFrac)
  train <- stocksDF[1:cutoff,]
  test <- stocksDF[(cutoff+1):nrow(stocksDF),]
  k <- findOptimalK(train)
  m <- mean(train$ratio); s <- sd(train$ratio)
  pos <- findPositions(test$ratio, m, s, k=k)
  
  if (plot) {
    plot(x=1:nrow(stocksDF), y=stocksDF$ratio,
         type = 'l', xlab = '', ylab = paste('Ratio (k=',k,')',sep = ''))
    abline(h = c(m-s*k, m, m+s*k), lty=c(2,1,2))
    abline(v = cutoff)
    for (p in pos) {
      points(x=p[1]+cutoff,y=stocksDF$ratio[p[1]+cutoff],col='green',pch=19)
      points(x=p[2]+cutoff,y=stocksDF$ratio[p[2]+cutoff],col='red',pch=19)
    }
  }
  
  return(positionProfit(test, pos))
}

simulateStockPair <- function(n=1000, sigma1=1, sigma2=1, rho=1, psi=0, b1=0, b2=0, plot=FALSE) {
  stock1 <- numeric(n); stock2 <- numeric(n)
  x1 <- 2; x2 <- 2
  for (t in 1:n) {
    stock1[t] <- 40 + b1*t + x1; stock2[t] <- 35 + b2*t + x2
    x1 <- rho*x1 + (1-rho)*psi*x2 + rnorm(1,0,sigma1)
    x2 <- rho*x2 + (1-rho)*psi*x1 + rnorm(1,0,sigma2)
  }
  r <- stock1/stock2
  df <- data.frame(stock1 = stock1, stock2 = stock2, ratio = r)
  if (plot) plotStocks(df)
  return(df)
}
simulateDistribution <- function(nrep = 100, returnCorrelation = FALSE, ...) {
  dist <- numeric(nrep)
  if (returnCorrelation) {
    for (i in 1:nrep) {
      df <- simulateStockPair(...)
      dist[i] <- cor(df$stock1, df$stock2)
    }
    return(dist)
  }
  for (i in 1:nrep) {
    dist[i] <- evaluatePairsTrading(simulateStockPair(...))
  }
  return(dist)
}
findPositionsExtension <- function(ratio, k=1) {
  len <- length(ratio)
  toReturn <- list(); lind <- 1
  finger <- 2
  openquestionmark <- FALSE
  m <- numeric(len); s <- numeric(len)
  for (i in 1:len) {
    m[i] <- mean(ratio[1:i])
    s[i] <- sd(ratio[1:i])
  }
  
  while (finger < len) {
    
    if (!openquestionmark) { #means we have to find the first place to open a position
      if (ratio[finger] > m[finger]+s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- 1
        openquestionmark <- TRUE
      }
      else if (ratio[finger] < m[finger]-s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- -1
        openquestionmark <- TRUE
      }
      else {
        finger <- finger + 1
      }
    }
    
    else { #find a place to close position
      if (highorlow == 1) {
        while (openquestionmark) {
          if (ratio[finger]<m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, 1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
      else {
        while(openquestionmark) {
          if (ratio[finger]>m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, -1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
    }
  }
  return(toReturn)
}
findPositionsExtension2 <- function(ratio, m, s, k=1) {
  len <- length(ratio)
  if (len <= 750) {
    stop('Not enough data')
  }
  toReturn <- list(); lind <- 1
  finger <- 750
  openquestionmark <- FALSE
  
  while (finger < len) {
    
    if (!openquestionmark) { #means we have to find the first place to open a position
      if (ratio[finger] > m[finger]+s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- 1
        openquestionmark <- TRUE
      }
      else if (ratio[finger] < m[finger]-s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- -1
        openquestionmark <- TRUE
      }
      else {
        finger <- finger + 1
      }
    }
    
    else { #find a place to close position
      if (highorlow == 1) {
        while (openquestionmark) {
          if (ratio[finger]<m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, 1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
      else {
        while(openquestionmark) {
          if (ratio[finger]>m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, -1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
    }
  }
  return(toReturn)
}
plotRatioExtension2 <- function(stockDF, k=1) {
  plotRatioExtension(stockDF, k=k)
  abline(v=750)
}
findOptimalKExtension <- function(stocksDF, plot=FALSE) {
  ratio <- stocksDF$ratio
  len <- length(ratio)
  m <- numeric(len); s <- numeric(len)
  for (i in 1:len) {
    m[i] <- mean(ratio[1:i])
    s[i] <- sd(ratio[1:i])
  }
  kmax <- max(abs(ratio[750:len]-m[750:len])/s[750:len])
  kvalues <- seq(0, kmax, length.out = 100)
  kprof <- sapply(kvalues, function(x) positionProfit(stocksDF = stocksDF,
                                                      findPositionsExtension2(ratio, m, s, k=x)))
  ind <- which(kprof == max(kprof))[1]
  bestk <- kvalues[ind]
  
  if (plot) {
    plot(x=kvalues, y=kprof, type='p',
         xlab='k value', ylab='Profit')
    points(x=bestk, y=kprof[ind], pch=19, col='red')
  }
  
  toReturn <- c(bestk, kprof[ind])
  return(toReturn)
}
findPositionsExtensionHelper <- function(ratio, m, s, k=1) {
  len <- length(ratio)
  toReturn <- list(); lind <- 1
  finger <- 1
  openquestionmark <- FALSE
  
  while (finger < len) {
    
    if (!openquestionmark) { #means we have to find the first place to open a position
      if (ratio[finger] > m[finger]+s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- 1
        openquestionmark <- TRUE
      }
      else if (ratio[finger] < m[finger]-s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- -1
        openquestionmark <- TRUE
      }
      else {
        finger <- finger + 1
      }
    }
    
    else { #find a place to close position
      if (highorlow == 1) {
        while (openquestionmark) {
          if (ratio[finger]<m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, 1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
      else {
        while(openquestionmark) {
          if (ratio[finger]>m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, -1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
    }
  }
  return(toReturn)
}
evaluatePairsTradingExtension <- function(stocksDF, trainingFrac = 0.5, plot = FALSE) {
  cutoff <- ceiling((nrow(stocksDF)-750)*trainingFrac)
  k <- findOptimalKExtension(stocksDF[1:(cutoff+750),])[1]
  
  ratio <- stocksDF$ratio; len <- length(ratio)
  m <- numeric(len); s <- numeric(len)
  for (i in 1:len) {
    m[i] <- mean(stocksDF$ratio[1:i])
    s[i] <- sd(stocksDF$ratio[1:i])
  }
  
  pos <- findPositionsExtensionHelper(ratio[(cutoff+751):len],
                                      m[(cutoff+751):len],
                                      s[(cutoff+751):len],
                                      k = k)
  
  if (plot) {
    plotRatioExtension2(stocksDF, k=k)
    for (p in pos) {
      points(x=p[1]+cutoff+750,y=stocksDF$ratio[p[1]+cutoff+750],col='green',pch=19)
      points(x=p[2]+cutoff+750,y=stocksDF$ratio[p[2]+cutoff+750],col='red',pch=19)
    }
    abline(v=cutoff+750, lty=2)
  }  
  
  test <- stocksDF[(cutoff+751):len,]
  return(positionProfit(test, pos))
}
```


The last thing we can do to find the effectiveness of the new strategy in comparision to the old one is to run a bunch of simulations and see how each strategy does on the same stock pair. After running this comparision on thousands of simulated pairs, no trend became very obvious. After creating a distribution of the difference in profit from the new strategy minus the profit from the old strategy, the distribution was always centered at zero. The issue was that outliers in the data set would consistently drag the mean either very high or very low so there was no discernable pattern to which strategy preforms better. Below is just an example mean of a distribution of 500 stock pairs.

```{r echo=FALSE}
simulateDistributionComparison <- function(nrep = 100, returnCorrelation = FALSE, ...) {
  dist <- numeric(nrep)
  if (returnCorrelation) {
    for (i in 1:nrep) {
      df <- simulateStockPair(...)
      dist[i] <- cor(df$stock1, df$stock2)
    }
    return(dist)
  }
  for (i in 1:nrep) {
    df <- simulateStockPair(...)
    dist[i] <- evaluatePairsTradingExtension(df) - evaluatePairsTrading(df)
  }
  return(dist)
}

dist <- simulateDistributionComparison(nrep=500, n = 2000)
message('The mean of the distribution is: ',mean(dist))
```

##Appendix of Code

```{r}
downloadStockPairDF <- function(stock1, stock2, start = 2010, nyears = 1) {
  stock1df <- downloadPriceDF(stock1, start = start, nyears = nyears)
  stock2df <- downloadPriceDF(stock2, start = start, nyears = nyears)
  if (!identical(stock1df$date, stock2df$date)) {
    stop('Dates are wack')
    #rid <- which(stock1df$date != stock2df$date)
    #stock1df <- stock1df[-rid, ]
    #stock2df <- stock2df[-rid, ]
  }
  df <- data.frame(stock1 = stock1df$price,
                   stock2 = stock2df$price,
                   ratio = stock1df$price/stock2df$price)
  return(df)
}

plotStocks <- function(stockPairDF) {
  max <- max(max(stockPairDF$stock1), max(stockPairDF$stock2))
  plot(x = 1:length(stockPairDF$stock1), y = stockPairDF$stock1, type = 'l', col = 'red',
       ylim = c(0, max), xlab = '', ylab = 'Price')
  lines(x = 1:length(stockPairDF$stock1), y = stockPairDF$stock2, col = 'blue')
  legend('topleft', legend = c('stock1', 'stock2'), col = c('red', 'blue'), lty = 1, cex = .8)
}
plotRatio <- function(stockDF, k=1) {
  m <- mean(stockDF$ratio); s <- sd(stockDF$ratio)
  plot(x=1:length(stockDF$ratio), y=stockDF$ratio,
       type='l',
       xlab='', ylab=paste('Ratio (k=',k,')', sep = ''))
  abline(h=c(m-k*s,m,m+k*s), lty=c(2,1,2))
}
findPositions <- function(ratio, m, s, k = 1) {
  len <- length(ratio)
  toReturn <- list(); lind <- 1
  finger <- 1
  openquestionmark <- FALSE
  
  while (finger < len) {

    if (!openquestionmark) { #means we have to find the first place to open a position
      highones <- which(ratio[finger:len]>(m+s*k))
      lowones <- which(ratio[finger:len]<(m-s*k))
      if (length(highones)>0 | length(lowones)>0) {
       start <- min(highones[1], lowones[1], na.rm = TRUE) + finger - 1
       highorlow <- ifelse(identical(start, highones[1]+finger-1),
                           1, -1)
       finger <- start
       openquestionmark <- TRUE
      }
      else finger <- len
    }
    else { #we need to find where to close the position
      close <- ifelse(highorlow == 1,
                      which(ratio[finger:len]<m)[1],
                      which(ratio[finger:len]>m)[1])
      if (is.na(close)) { #means there is no good close; must close on last day
        close <- len
      }
      else close <- close+finger-1
      
      finger <- close
      toReturn[[lind]] <- c(start, close, highorlow)
      lind <- lind + 1
      openquestionmark <- FALSE
    }
       
  }#end of while loop
  return(toReturn)
}
addPositions <- function(ratio, positions) {
  for (p in positions) {
    points(x=p[1],y=ratio[p[1]],col='green',pch=19)
    points(x=p[2],y=ratio[p[2]],col='red',pch=19)
  }
}
positionProfit <- function(stocksDF, positions, net = TRUE) {
  if (length(positions) == 0) return(0)
  cashcash <- numeric(length(positions))
  i <- 1
  for (p in positions) {
    shares1 <- 1/stocksDF$stock1[p[1]]
    shares2 <- 1/stocksDF$stock2[p[1]]
    profit1 <- p[3]*-1 * shares1 * stocksDF$stock1[p[2]]
    profit2 <- p[3] * shares2 * stocksDF$stock2[p[2]]
    fees <- 0.003 * (1+1+abs(profit1)+abs(profit2))
    cashcash[i] <- profit1+profit2-fees
    i <- i+1
  }
  if (net) return(sum(cashcash))
  return(cashcash)
}
findOptimalK <- function(stocksDF, plot = FALSE) {
  ratio <- stocksDF$ratio; m <- mean(stocksDF$ratio); s <- sd(stocksDF$ratio)
  kmax <- max(abs(ratio - m))/s
  kvalues <- seq(0, kmax, length = 100)
  kprof <- sapply(kvalues, function(x) positionProfit(stocksDF = stocksDF,
                                                      findPositions(ratio, m, s, k=x)))
  ind <- which(kprof == max(kprof))[1]
  bestk <- kvalues[ind]
  
  if (plot) {
    plot(x=kvalues, y=kprof, type='p',
         xlab='k value', ylab='Profit')
    points(x=bestk, y=kprof[ind], pch=19, col='red')
  }
  
  return(bestk)
}
evaluatePairsTrading <- function(stocksDF, trainingFrac = 0.5, plot = FALSE) {
  cutoff <- ceiling(nrow(stocksDF)*trainingFrac)
  train <- stocksDF[1:cutoff,]
  test <- stocksDF[(cutoff+1):nrow(stocksDF),]
  k <- findOptimalK(train)
  m <- mean(train$ratio); s <- sd(train$ratio)
  pos <- findPositions(test$ratio, m, s, k=k)
  
  if (plot) {
    plot(x=1:nrow(stocksDF), y=stocksDF$ratio,
         type = 'l', xlab = '', ylab = paste('Ratio (k=',k,')',sep = ''))
    abline(h = c(m-s*k, m, m+s*k), lty=c(2,1,2))
    abline(v = cutoff)
    for (p in pos) {
      points(x=p[1]+cutoff,y=stocksDF$ratio[p[1]+cutoff],col='green',pch=19)
      points(x=p[2]+cutoff,y=stocksDF$ratio[p[2]+cutoff],col='red',pch=19)
    }
  }
  
  return(positionProfit(test, pos))
}

simulateStockPair <- function(n=1000, sigma1=1, sigma2=1, rho=1, psi=0, b1=0, b2=0, plot=FALSE) {
  stock1 <- numeric(n); stock2 <- numeric(n)
  x1 <- 2; x2 <- 2
  for (t in 1:n) {
    stock1[t] <- 40 + b1*t + x1; stock2[t] <- 35 + b2*t + x2
    x1 <- rho*x1 + (1-rho)*psi*x2 + rnorm(1,0,sigma1)
    x2 <- rho*x2 + (1-rho)*psi*x1 + rnorm(1,0,sigma2)
  }
  r <- stock1/stock2
  df <- data.frame(stock1 = stock1, stock2 = stock2, ratio = r)
  if (plot) plotStocks(df)
  return(df)
}
simulateDistribution <- function(nrep = 100, returnCorrelation = FALSE, ...) {
  dist <- numeric(nrep)
  if (returnCorrelation) {
    for (i in 1:nrep) {
      df <- simulateStockPair(...)
      dist[i] <- cor(df$stock1, df$stock2)
    }
    return(dist)
  }
  for (i in 1:nrep) {
    dist[i] <- evaluatePairsTrading(simulateStockPair(...))
  }
  return(dist)
}
findPositionsExtension <- function(ratio, k=1) {
  len <- length(ratio)
  toReturn <- list(); lind <- 1
  finger <- 2
  openquestionmark <- FALSE
  m <- numeric(len); s <- numeric(len)
  for (i in 1:len) {
    m[i] <- mean(ratio[1:i])
    s[i] <- sd(ratio[1:i])
  }
  
  while (finger < len) {
    
    if (!openquestionmark) { #means we have to find the first place to open a position
      if (ratio[finger] > m[finger]+s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- 1
        openquestionmark <- TRUE
      }
      else if (ratio[finger] < m[finger]-s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- -1
        openquestionmark <- TRUE
      }
      else {
        finger <- finger + 1
      }
    }
    
    else { #find a place to close position
      if (highorlow == 1) {
        while (openquestionmark) {
          if (ratio[finger]<m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, 1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
      else {
        while(openquestionmark) {
          if (ratio[finger]>m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, -1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
    }
  }
  return(toReturn)
}
findPositionsExtension2 <- function(ratio, m, s, k=1) {
  len <- length(ratio)
  if (len <= 750) {
    stop('Not enough data')
  }
  toReturn <- list(); lind <- 1
  finger <- 750
  openquestionmark <- FALSE
  
  while (finger < len) {
    
    if (!openquestionmark) { #means we have to find the first place to open a position
      if (ratio[finger] > m[finger]+s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- 1
        openquestionmark <- TRUE
      }
      else if (ratio[finger] < m[finger]-s[finger]*k) {
        start <- finger
        finger <- finger + 1
        highorlow <- -1
        openquestionmark <- TRUE
      }
      else {
        finger <- finger + 1
      }
    }
    
    else { #find a place to close position
      if (highorlow == 1) {
        while (openquestionmark) {
          if (ratio[finger]<m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, 1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
      else {
        while(openquestionmark) {
          if (ratio[finger]>m[finger]) {
            toReturn[[lind]] <- c(start, finger, 1)
            lind <- lind + 1
            openquestionmark <- FALSE
          }
          else if (finger == len) {
            toReturn[[lind]] <- c(start, finger, -1)
            openquestionmark <- FALSE
          }
          else {
            finger <- finger + 1
          }
        }
      }
    }
  }
  return(toReturn)
}
plotRatioExtension2 <- function(stockDF, k=1) {
  plotRatioExtension(stockDF, k=k)
  abline(v=750)
}
findOptimalKExtension <- function(stocksDF, plot=FALSE) {
  ratio <- stocksDF$ratio
  len <- length(ratio)
  m <- numeric(len); s <- numeric(len)
  for (i in 1:len) {
    m[i] <- mean(ratio[1:i])
    s[i] <- sd(ratio[1:i])
  }
  kmax <- max(abs(ratio[750:len]-m[750:len])/s[750:len])
  kvalues <- seq(0, kmax, length.out = 100)
  kprof <- sapply(kvalues, function(x) positionProfit(stocksDF = stocksDF,
                                                      findPositionsExtension2(ratio, m, s, k=x)))
  ind <- which(kprof == max(kprof))[1]
  bestk <- kvalues[ind]
  
  if (plot) {
    plot(x=kvalues, y=kprof, type='p',
         xlab='k value', ylab='Profit')
    points(x=bestk, y=kprof[ind], pch=19, col='red')
  }
  
  toReturn <- c(bestk, kprof[ind])
  return(toReturn)
}
evaluatePairsTradingExtension <- function(stocksDF, trainingFrac = 0.5, plot = FALSE) {
  cutoff <- ceiling((nrow(stocksDF)-750)*trainingFrac)
  k <- findOptimalKExtension(stocksDF[1:(cutoff+750),])[1]
  
  ratio <- stocksDF$ratio; len <- length(ratio)
  m <- numeric(len); s <- numeric(len)
  for (i in 1:len) {
    m[i] <- mean(stocksDF$ratio[1:i])
    s[i] <- sd(stocksDF$ratio[1:i])
  }
  
  pos <- findPositionsExtensionHelper(ratio[(cutoff+751):len],
                                      m[(cutoff+751):len],
                                      s[(cutoff+751):len],
                                      k = k)
  
  if (plot) {
    plotRatioExtension2(stocksDF, k=k)
    for (p in pos) {
      points(x=p[1]+cutoff+750,y=stocksDF$ratio[p[1]+cutoff+750],col='green',pch=19)
      points(x=p[2]+cutoff+750,y=stocksDF$ratio[p[2]+cutoff+750],col='red',pch=19)
    }
    abline(v=cutoff+750, lty=2)
  }  
  
  test <- stocksDF[(cutoff+751):len,]
  return(positionProfit(test, pos))
}
simulateDistributionComparison <- function(nrep = 100, returnCorrelation = FALSE, ...) {
  dist <- numeric(nrep)
  if (returnCorrelation) {
    for (i in 1:nrep) {
      df <- simulateStockPair(...)
      dist[i] <- cor(df$stock1, df$stock2)
    }
    return(dist)
  }
  for (i in 1:nrep) {
    df <- simulateStockPair(...)
    dist[i] <- evaluatePairsTradingExtension(df) - evaluatePairsTrading(df)
  }
  return(dist)
}
```